Trends from 30-year observations of downward solar irradiance in Thessaloniki, Greece

### 2023-11-29 MDPI_submission.Rmd
Commit 2023-11-29 22:14

LT$cor.df           <- NULL
LT$cor.null_value   <- NULL
LT$cor.alternative  <- NULL
LT$cor.method       <- NULL
LT$cor.data_name    <- NULL
LT$cor.p            <- NULL
LT$conf.level       <- NULL

@@

## numeric precision of table
LT$slope    <- signif(LT$slope,   3)
LT$slope.p  <- signif(LT$slope.p, 3)
## rename sky conditions
LT$DATA <- sub(" cond.", "", LT$DATA)

@@

```{r trendtable, echo=F}
LT2$`Sky conditions` <- sub(" sky", " skies", LT2$`Sky conditions`)
x_latex <- knitr::kable(LT2,
                        format      = "latex",
                        booktabs    = TRUE,
                        escape      = TRUE,
                        format.args = list(digits = 4),
                        linesep     = "",
                        align       = "crrr",
                        centering   = FALSE,
                        table.envir = "table",
                        position    = "H",
                        caption     = paste0("Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))

@@

spp$cor.df             <- NULL
spp$cor.null_value     <- NULL
spp$cor.alternative    <- NULL
spp$cor.method         <- NULL
spp$cor.data_name      <- NULL
spp$cor.p              <- NULL
spp$conf.level         <- NULL

@@

@@

Similarly to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
(Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
For all-sky conditions the trend in SDR in winter is the largest
($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$),
followed by the trend in autumn
($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$,
a value close to the long term trend) both statistically significant above the $99\,\%$ confidence level.
In spring and summer, the trends are much smaller and of lesser statistical significance.
These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
Moreover, the trends under clear- and cloudy-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.

@@

<!-- \footnotesize -->

```{r trendseasontable, echo=F}
## arrange columns
# spp <- spp[,c(
#       "Sky condition",
#       "Season",
#       "Trend [%/year]",
#       # "Trend p-value",
#       "Statistical signif. [%]"
#       )]

@@

```
<!-- \normalsize -->

@@

The observed brightening on SDR over Thessaloniki is dependent on SZA (larger SZAs lead to stronger brightening).
The trend is also dependent on season, with winter showing the strongest statistically significant trend of
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All sky"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
for all- and clear-skies, respectively, in contrast to spring and summer.
The trends for autumn are also significant but smaller (
$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All sky"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Clear sky"], 2)`\,\%/y$
for all- and clear-skies, respectively).
Our findings are in agreement with other studies for the region.

>>>>

LT$cor.df           <- NULL
LT$cor.null_value   <- NULL
LT$cor.alternative  <- NULL
LT$cor.method       <- NULL
LT$cor.data_name    <- NULL
LT$cor.p            <- NULL
LT$conf.level       <- NULL
LT$cor.conf_int_1   <- NULL
LT$cor.conf_int_2   <- NULL

@@

## numeric precision of table
LT$slope    <- signif(LT$slope,   3)
LT$slope.p  <- signif(LT$slope.p, 3)
## rename sky conditions
LT$DATA <- sub(" cond.", "", LT$DATA)
LT$DATA <- sub(" sky", " skies", LT$DATA)

@@

```{r trendtable, echo=F}

x_latex <- knitr::kable(LT2,
                        format      = "latex",
                        booktabs    = TRUE,
                        escape      = TRUE,
                        format.args = list(digits = 4),
                        linesep     = "",
                        align       = "crrr",
                        centering   = FALSE,
                        table.envir = "table",
                        position    = "H",
                        caption     = paste0("Trends in SDR daily means for different sky conditions for the period ", year(min(ALL_1_daily_DESEAS$Date)), " - ", year(max(ALL_1_daily_DESEAS$Date)), "."))

@@

spp$cor.df             <- NULL
spp$cor.null_value     <- NULL
spp$cor.alternative    <- NULL
spp$cor.method         <- NULL
spp$cor.data_name      <- NULL
spp$cor.p              <- NULL
spp$conf.level         <- NULL
spp$cor.conf_int_1     <- NULL
spp$cor.conf_int_2     <- NULL

@@

spp <- rename(spp, "Trend [%/year]" = slope)
spp <- rename(spp, "Sky condition"  = DATA )

@@

Similarly to the long term trends discussed above, we have calculated the trend of the anomalies for the three different sky conditions, and for each season of the year, using the corresponding mean monthly values
(Figure\ \ref{fig:seasonalALL} and Table\ \ref{tab:trendseasontable}).
For all-sky conditions the trend in SDR in winter is the largest
($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$),
followed by the trend in autumn
($`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$,
a value close to the long term trend) both statistically significant above the $99\,\%$ confidence level.
In spring and summer, the trends are much smaller and of lesser statistical significance.
These seasonal differences indicate a possible relation of the trends in SDR to trends of clouds during winter and autumn.
For clear-skies, the trend in winter is $`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$, larger than for all-skies ($`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`\,\%/y$), which is another indication of a decreasing trend in cloud optical thickness.
Moreover, the trends under clear- and cloudy-sky conditions are almost complementary to each other, particularly for winter and autumn, where the signal is stronger.
During spring and summer the statistical significance is very low and the actual trend too small for a meaningful comparison.

@@

```{r trendseasontable, echo=F}

## arrange columns
# spp <- spp[,c(
#       "Sky condition",
#       "Season",
#       "Trend [%/year]",
#       # "Trend p-value",
#       "Statistical signif. [%]"
#       )]

@@

@@

The observed brightening on SDR over Thessaloniki is dependent on SZA (larger SZAs lead to stronger brightening).
The trend is also dependent on season, with winter showing the strongest statistically significant trend of
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "All skies"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Winter" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$
for all- and clear-skies, respectively, in contrast to spring and summer.
The trends for autumn are also significant but smaller (
$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "All skies"], 2)`$
and
$`r signif(spp$"Trend [%/year]"[spp$Season == "Autumn" & spp$"Sky condition" == "Clear skies"], 2)`\,\%/y$
for all- and clear-skies, respectively).
Our findings are in agreement with other studies for the region.

### 2023-11-29 MDPI_submission.tex
Commit 2023-11-29 22:14

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky conditions & Trend [\%/year] & Statistical signif. [\%] & Days with data\\
\midrule
All skies & 0.376 & 100.00 & 10256\\
Clear skies & 0.349 & 100.00 & 10256\\
Cloudy skies & -0.276 & 99.97 & 5067\\
\bottomrule
\end{tabu}
\end{table}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X}
\toprule
Sky condition & Season & Trend [\%/year] & Statistical signif. [\%]\\
\midrule
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{99.9}\\

 & Spring & 0.1450 & 81.8\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{89.4}\\

\multirow{-4}{*}{\centering\arraybackslash All skies} & Autumn & 0.4310 & 99.3\\
\cmidrule{1-4}
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{100.0}\\

 & Spring & 0.0613 & 38.6\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{25.9}\\

\multirow{-4}{*}{\centering\arraybackslash Clear skies} & Autumn & 0.3670 & 97.2\\
\cmidrule{1-4}
\cellcolor{gray!6}{} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{98.9}\\

 & Spring & -0.0991 & 37.2\\

\cellcolor{gray!6}{} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{21.5}\\

\multirow{-4}{*}{\centering\arraybackslash Cloudy skies} & Autumn & -0.4000 & 89.5\\
\bottomrule
\end{tabu}
\end{table}

>>>>

\caption{\label{tab:trendtable}Trends in SDR daily means for different sky conditions for the period 1993 - 2023.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\raggedleft}X>{\raggedleft}X}
\toprule
slope & slope.p & Rsqrd & RsqrdAdj & cor.t & cor.esimate & cor.conf\_int\_1 & cor.conf\_int\_2 & DATA & N & slope.stat\_sig\\
\midrule
0.376 & 0.00000 & 0.008308 & 0.008211 & 9.268 & 0.09115 & 0.07192 & 0.11031 & All skies & 10256 & 100.00\\
0.349 & 0.00000 & 0.006183 & 0.006087 & 7.987 & 0.07864 & 0.05937 & 0.09784 & Clear skies & 10256 & 100.00\\
-0.276 & 0.00026 & 0.002631 & 0.002434 & -3.655 & -0.05129 & -0.07872 & -0.02379 & Cloudy skies & 5067 & 99.97\\
\bottomrule
\end{tabu}
\end{table}

@@

\caption{\label{tab:trendseasontable}SDR linear trends of monthly anomalies for each season of the year.}
\begin{tabu} to \linewidth {>{\centering\arraybackslash}p{8em}>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X>{\centering}X>{\raggedleft}X>{\raggedleft}X>{\centering}X}
\toprule
Trend [\%/year] & slope.p & cor.t & cor.esimate & cor.conf\_int\_1 & cor.conf\_int\_2 & Sky condition & Season & slope.stat\_sig\\
\midrule
\cellcolor{gray!6}{0.6860} & \cellcolor{gray!6}{0.001300} & \cellcolor{gray!6}{3.5748458} & \cellcolor{gray!6}{0.5598043} & \cellcolor{gray!6}{0.2499438} & \cellcolor{gray!6}{0.7656559} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{99.9}\\
\cmidrule{1-9}
0.1450 & 0.182000 & 1.3668258 & 0.2460126 & -0.1186724 & 0.5522150 & All skies & Spring & 81.8\\
\cmidrule{1-9}
\cellcolor{gray!6}{0.1200} & \cellcolor{gray!6}{0.106000} & \cellcolor{gray!6}{1.6689762} & \cellcolor{gray!6}{0.3007995} & \cellcolor{gray!6}{-0.0666976} & \cellcolor{gray!6}{0.5964339} & \cellcolor{gray!6}{All skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{89.4}\\
\cmidrule{1-9}
0.4310 & 0.006650 & 2.9317564 & 0.4846362 & 0.1506747 & 0.7193136 & All skies & Autumn & 99.3\\
\cmidrule{1-9}
\cellcolor{gray!6}{0.8260} & \cellcolor{gray!6}{0.000259} & \cellcolor{gray!6}{4.1794536} & \cellcolor{gray!6}{0.6198226} & \cellcolor{gray!6}{0.3341758} & \cellcolor{gray!6}{0.8011848} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{100.0}\\
\cmidrule{1-9}
0.0613 & 0.614000 & 0.5095289 & 0.0941964 & -0.2691265 & 0.4340490 & Clear skies & Spring & 38.6\\
\cmidrule{1-9}
\cellcolor{gray!6}{-0.0307} & \cellcolor{gray!6}{0.741000} & \cellcolor{gray!6}{-0.3343454} & \cellcolor{gray!6}{-0.0630596} & \cellcolor{gray!6}{-0.4139251} & \cellcolor{gray!6}{0.3041187} & \cellcolor{gray!6}{Clear skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{25.9}\\
\cmidrule{1-9}
0.3670 & 0.028300 & 2.3126418 & 0.4004713 & 0.0469803 & 0.6648218 & Clear skies & Autumn & 97.2\\
\cmidrule{1-9}
\cellcolor{gray!6}{-0.8820} & \cellcolor{gray!6}{0.011100} & \cellcolor{gray!6}{-2.7197359} & \cellcolor{gray!6}{-0.4571342} & \cellcolor{gray!6}{-0.7018198} & \cellcolor{gray!6}{-0.1159631} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Winter} & \cellcolor{gray!6}{98.9}\\
\cmidrule{1-9}
-0.0991 & 0.628000 & -0.4892643 & -0.0904814 & -0.4310031 & 0.2725985 & Cloudy skies & Spring & 37.2\\
\cmidrule{1-9}
\cellcolor{gray!6}{0.0444} & \cellcolor{gray!6}{0.785000} & \cellcolor{gray!6}{0.2750824} & \cellcolor{gray!6}{0.0519156} & \cellcolor{gray!6}{-0.3142309} & \cellcolor{gray!6}{0.4046170} & \cellcolor{gray!6}{Cloudy skies} & \cellcolor{gray!6}{Summer} & \cellcolor{gray!6}{21.5}\\
\cmidrule{1-9}
-0.4000 & 0.105000 & -1.6769099 & -0.3020992 & -0.5973542 & 0.0652743 & Cloudy skies & Autumn & 89.5\\
\bottomrule
\end{tabu}
\end{table}
